---
title: "Working with XML and JSON in R"
author: "Srini Illapani"
date: "October 18, 2015"
output: html_document
---

## 1. Loading and parsing XML file

1. Loading required packages

```{r, eval=TRUE}
library(XML)
library(RCurl)
```

2. XML package does not support https. So we will get the file from the secured site
using RCurl first and then use xmlTreeParse(). 

```{r, eval=TRUE}
# xml file url
xml_file = getURL("https://raw.githubusercontent.com/isrini/SI_IS607/master/books.xml")

# parse html content
xml_file2 = xmlTreeParse(xml_file)
class(xml_file2)

head(xml_file2)

```


## 2. Loading and parsing HTML file

1. Loading required packages

```{r, eval=FALSE}
library(XML)
library(RCurl)
```

2. XML package does not support https. So we will get the file from the secured site
using RCurl first and then use htmlTreeParse().  

```{r, eval=TRUE}
# html file url
html_file = getURL("https://raw.githubusercontent.com/isrini/SI_IS607/master/books.html")

# parse html content
html_file2 = htmlTreeParse(html_file)
class(html_file2)

head(html_file2)

```



## 3. Loading and parsing JSON file

1. Loading required packages

```{r, eval=TRUE}
library(RJSONIO)
```

2. Get the file from the secured site using RCurl first and then use fromJSON().  

```{r, eval=TRUE}
# html file url
json_file = getURL("https://raw.githubusercontent.com/isrini/SI_IS607/master/books.json")

# fromJSON() in package RJSONIO
json_file2 = RJSONIO::fromJSON(json_file)

# class
class(json_file2)

head(json_file2)
```


## Conclusion:

#### Based on the outputs from the three data frames as seen above, the data frames are not identical.  
  
  


