---
title: "Illapani_wk11"
author: "Srini Illapani"
date: "November 22, 2015"
output: html_document
---

## Document Classification



### Load the two files into R:

The dataset is based on the real data examples of spam emails at - https://archive.ics.uci.edu/ml/datasets/Spambase
The dataset contains 57 attributes. Each attribute refers to the occurance or frequency of a specific word in the
spambase email files.  

It also contains attributes which show the number of certain chars in the email like ";", "$", "!" etc.  

And the last attribute denotes whether the email was considered spam (1) or ham (not spam) (0).  


```{r, eval=TRUE}
# Reading the file from my repository into R
file = "https://raw.githubusercontent.com/isrini/SI_IS607/master/data.csv"
data <- read.csv(file, header=TRUE, sep="|")
```



```{r, eval=TRUE}
# Make the last column 'y' (spam yes or no) a factor variable for binary classification
data$y <- as.factor(data$y)

# Sample of 100 rows
data <- data[sample(nrow(data), 500),]
colnames(data)

```


### Load the library packages

```{r, eval=TRUE}
library(caret)
library(rpart)
library(e1071)
```



### Split the data in to two - train and test

```{r, eval=TRUE}
# split data into a train and test set
i <- 1:nrow(data)
trainIndex <- sample(i, trunc(length(i)/4))

data_test <- data[trainIndex,]
data_train  <- data[-trainIndex,]
```



### Create the SVM and rpart model

```{r, eval=TRUE}
# Both for the SVM and the partitioning tree (via rpart()), we fit the model and try to predict the test set values:
# create a model using the training data set
svm.model <- svm(y~., data = data_train)
rpart.model <- rpart(y~., data = data_train)

```


### Evaluate the model for prediction 

```{r, eval=TRUE}

svm.pred <- predict(svm.model,data_test[,-58])
rpart.pred <- predict(rpart.model,data_test[,-58], type = "class")


```

### A cross-tabulation of the true versus the predicted values yields.

```{r, eval=TRUE}
# compute svm confusion matrix
svm.cm <- table(pred = svm.pred, true = data_test[,58])
svm.cm

# compute rpart confusion matrix
rpart.cm <- table(pred = rpart.pred, true = data_test[,58])
rpart.cm

# compare the performance of the two methods by computing the respective accuracy rates and the kappa indices

classAgreement(svm.cm)
classAgreement(rpart.cm)

```

The results are almost same for SVM and rpart models.






