---
title: "IS 607 - Project 2"
author: "Srini Illapani"
date: "October 10, 2015"
output: html_document
---
## Wide data set 1 - Source fivethirtyeight  
The source https://github.com/fivethirtyeight/data/blob/master/forecast-methodology/historical-senate-predictions.csv).

The data depicts the win/loss probability predictions for the years 2008 to 2012 US Senate candidates by state.


#### 1. Load the required packages:

```{r, eval=TRUE, warning=FALSE, message=FALSE}
library(dplyr)
library(tidyr)
library(knitr)
library(ggplot2)
```

#### 2. After downloding the source file from github, read the CSV file into an R object

```{r, eval=TRUE, warning=FALSE}
senate_pred <- read.csv("https://raw.githubusercontent.com/isrini/SI_IS607/master/historical-senate-predictions.csv")
glimpse(senate_pred)
```



#### 3. Perform data wrangling with tidyr and dplyr packages

The source file has data from the years 2008 to 2012 and does not reflect all the 50 states for each of those years.
For the purpose of this study, lets filter the records for the year 2012 Senate race.

```{r, eval=TRUE, warning=FALSE}
senate_2012 <- filter(senate_pred, year == '2012')
senate_2012 <- select(senate_2012, state, candidate, forecast_prob, result, winflag)

head(senate_2012)
```

The data set is now in a wide format that would still need some tidying and formatting before we can analyze the final set.
Renaming the data from the columns 'result' and 'winflag'

```{r, eval=TRUE, warning=FALSE}
senate_2012$result <- gsub("Win", "Winner", senate_2012$result)
senate_2012$result <- gsub("Loss", "Loser", senate_2012$result)

senate_2012$winflag <- gsub('1', "win_forecast_prob", senate_2012$winflag)
senate_2012$winflag <- gsub('0', "loss_forecast_prob", senate_2012$winflag)

head(senate_2012)
```

Now, spread the column 'result' in to two columns to show the 'loser' and 'winner' by name.  

```{r, eval=TRUE, warning=FALSE}
senate_2012 <- spread(senate_2012, result, candidate)
head(senate_2012)
```

Do the same with the 'winflag' column and show the forecast probability as losss and win probability columns.  


```{r, eval=TRUE, warning=FALSE}
senate_2012 <- spread(senate_2012, winflag, forecast_prob)
head(senate_2012)
```

We have the State row occuring twice to show each candidate and the corresponding data. Using the group by
function and summarise, the two State rows have been merged in to one row. And the NA references removed.

```{r, eval=TRUE, warning=FALSE}
senate_2012 <- senate_2012 %>%
      group_by(state) %>%
      summarise_each(funs(first(.[!is.na(.)])))
    
head(senate_2012)
```


### Here is the transformed Tidy data set

```{r, eval=TRUE}
kable(senate_2012)
```

### 3. Perform win and loss probability analysis.
  
We will filter for those with less than 100% win or loss accuracy prediction, to analyze the others who were not 
in an one sided race.

```{r, eval=TRUE}
senate_2012_winners <- select(senate_2012, state, Winner, win_forecast_prob)

senate_2012_winners %>%
  mutate(win_accuracy = win_forecast_prob * 100 ) %>%
  filter(win_accuracy < 100) %>%
  select(state, Winner, win_accuracy)
```


```{r, eval=TRUE}
senate_2012_losers <- select(senate_2012, state, Loser, loss_forecast_prob)

senate_2012_losers %>%
  mutate(loss_accuracy = loss_forecast_prob * 100 ) %>%
  filter(loss_accuracy < 100) %>%
  select(state, Loser, loss_accuracy)
```

### 4. Visual Plotting of the winning and losing predictions


```{r, eval=TRUE}
library(ggplot2)
qplot(win_forecast_prob, data = senate_2012_winners, geom = "dotplot")
```

Almost all the win predictions came true except for one senate race, the single dot to the left. This candidate
won the race even though the prediction was very low probability for win. 

```{r, eval=TRUE}
library(ggplot2)
qplot(loss_forecast_prob, data = senate_2012_losers, geom = "dotplot")
```

This is a similar plot with the reverse view of the win_prediction probability for the senate race losers. The one senate
race that we saw above is also shown here as a wrong prediction. The person who was predicted to win lost and the single dot 
to the right represents that.  

Overall, This is a high quality predictive analysis by FiveThirtyEight.com based on the poll data from various polling sources
and other indicators or variables.  


***  
## Wide data set 2 (PanTHERIA - Suggested By Veneranda Skrelja )  

Jones, K.E., et al. PanTHERIA: a species-level database of life history, ecology, and geography of extant and recently 
extinct mammals. Ecology 90:2648. http://esapubs.org/archive/ecol/E090/184/

```{r, eval=TRUE}
pantheria <-  "http://esapubs.org/archive/ecol/E090/184/PanTHERIA_1-0_WR05_Aug2008.txt"
download.file(pantheria, destfile = "mammals.txt")
```

```{r, eval=TRUE}

# Read the data from the downloaded txt file in to a data set called 'mammals'
mammals <- read.table("mammals.txt", sep = "\t", header = TRUE,  stringsAsFactors = FALSE)

# Tidy the column names
names(mammals) <- sub("X[0-9._]+", "", names(mammals))
names(mammals) <- sub("MSW05_", "", names(mammals))

# Select the columns names to be used for further transformation
mammals <- dplyr::select(mammals, Order, Binomial, AdultBodyMass_g, AdultHeadBodyLen_mm, HomeRange_km2, LitterSize)
head(mammals)

# Renaming the column names to lower case and adding '_' in between words and other formatting
names(mammals) <- gsub("([A-Z])", "_\\L\\1", names(mammals), perl = TRUE)
names(mammals) <- gsub("^_", "", names(mammals), perl = TRUE)
mammals[mammals == -999] <- NA
names(mammals)[names(mammals) == "binomial"] <- "species"

head(mammals)
```

### 1. Data manipulation

```{r, eval=TRUE}
# Mutate - to add new columns using current data
head(mutate(mammals, adult_body_mass_kg = adult_body_mass_g / 1000))

# Summarising data by grouping
head(summarise(group_by(mammals, order),
  mean_mass = mean(adult_body_mass_g, na.rm = TRUE)))

# Piping data
piping = mammals %>%
    mutate(mass_to_length = adult_body_mass_g / adult_head_body_len_mm) %>%
    arrange(desc(mass_to_length)) %>%
    select(species, mass_to_length, litter_size)

head(piping)
```

### 2. Data Analysis

```{r, eval=TRUE}
# Using group by and summarise to find the mean litter_size by species
mean = group_by(mammals,species) %>% 
    filter(litter_size != 'NA') %>% 
    summarise(mean(litter_size))

head(mean)
```


