---
title: "IS607 - Week3"
author: "Srini Illapani"
date: "September 9, 2015"
output: html_document
---

####1. For this assignment, I downloaded the Pittsburgh bridges dataset from
https://archive.ics.uci.edu/ml/datasets/Pittsburgh+Bridges.
Uploaded the file to my github repository and used the curl package to retrieve the dataset from my repository.

```{r, eval = TRUE}
library(curl)
bridges <- read.csv(curl("https://raw.githubusercontent.com/isrini/SI_IS607/master/bridges.data.version1"), header = FALSE)
head(bridges)

```


####2. The data set includes 108 rows and 13 columns. It does not have the headers to describe the columns.
Lets add the column names or headers to the data set using the following code and then displaying the data using a function from knitr package.

```{r, eval=TRUE}
colnames(bridges) <- c("identifier", "river", "location", "erected","purpose", "length", "lanes", "clear-g", "T_or_D", "material", "span", "rel_l", "type")

rownames(bridges) <- NULL
library(knitr)
kable(head(bridges, 20))
```


####3. Data transformation:
To transform column 'river' data to their full abbreviation values, the below function is being used.

```{r, eval=TRUE}
bridges$river <- ifelse(bridges$river=="A", "Allegheny", 
              ifelse(bridges$river=="M", "Monongahela",
              ifelse(bridges$river=="O", "Ohio", "N/A")
                     
))

```


####4. Data transformation:
To transform column 'purpose' data for 'RR' to its full abbreviation, the below code is being used.

```{r, eval=TRUE}

bridges$purpose <- gsub("RR", "RAILROAD", bridges$purpose)
       
```


####5. View the transformed data

```{r, eval=TRUE}
kable(head(bridges, 20))
```

