---
title: "IS607_Project 1"
author: "Srini Illapani"
date: "September 20, 2015"
output: html_document
---
***

#### Question:

Use the provided text file with chess tournament results where the information has some structure.
Your job is to create an R Markdown file that generates a .CSV file (that could for example be imported into a SQL database)
with the following information for all of the players:

Player's Name, Player's State, Total Number of Points, Player's Pre-Rating, and Average Pre Chess Rating of Opponents

***

1. First, import chess data from the txt file into R:


```{r, eval = TRUE, warning=FALSE}
score_card = readLines("https://raw.githubusercontent.com/isrini/SI_IS607/master/tournamentinfo.txt");
```



2. Using the stringr library to extract the data:

```{r, eval = TRUE, warning=FALSE}
library(stringr);
library(knitr);
```


3. Thw txt file has two rows of data for each player. Lets extract each row separately and remove the blanks - space and tab:

```{r, eval = TRUE}
row1 = unlist(str_extract_all(score_card,"^[[:blank:]]+\\d{1,2}.+"))
row2 = unlist(str_extract_all(score_card,"^[[:blank:]]+[A-Z]{2}.+"))
head(row1);
head(row2);
```


4. And extract the chess player name from row 1, the state they represent from row 2 and total points from row 1

```{r, eval = TRUE}
# names are in upper case
player_name = unlist(str_extract_all(row1,"(\\b[[:upper:]-]+\\b\\s)+(\\b[[:upper:]-]+\\b){1}"))

# state name is also in upper case
state_name = unlist(str_extract_all(row2,"[[:upper:]]{2}" ))

# points are decimal numbers
points = as.numeric(unlist(str_extract_all(row1,"\\d(.)\\d")))                                  
```


5. Now lets extract the pre rating data from row 2:

```{r, eval = TRUE}
# some of the ratings have a 'P' followedd by additional numbers, we do not need this info, replace it
pre_rating <- str_replace_all(row2,pattern="[P]\\d{1,}"," ")

# removing the prefix of R: from the pre_rating
r_colon <- str_extract_all(pre_rating,"([R(:)][[:blank:]]+\\d{3,}+)") 

# the pre rating and converting it to a numeric
pre_rating <- as.numeric(str_extract_all(r_colon,"\\d{3,}"))          

```


6. Lets put all these together in a data frame:

```{r, eval = TRUE}
score_df <- data.frame(player_name, state_name, points, pre_rating);
head(score_df);
```


7. Now lets get the opponent player numbers for each of the 64 players from row 1:

```{r, eval = TRUE}
# each player number has a prefix of W,L or D followed by a space, we do not need this
player_opponents <- str_extract_all(row1,"[WLD][[:blank:]]+\\d{1,2}") 

Opponents <-str_extract_all(player_opponents,"\\d{1,2}")

# convert the player numbers to numeric
opponents <- lapply(Opponents, as.numeric)                            
```


8. We need to calculate the average pre chess rating for each player based on their opponents pre_rating numbers. Lets create a function to calculate the
average rating using their opponents information from above.

```{r, eval = TRUE}
avg_opp_rate <- function(a){
  
  rating <- 0
  no_of_opponents <- length(a)

  for (i in a){
    rating <- rating + score_df[i,"pre_rating"]}
    return(rating / no_of_opponents)
  }

avg_pre_chess_rating <- unlist(lapply(opponents, avg_opp_rate));

# convert to numeric as well as rounding off the ranking
avg_pre_chess_rating <- round(avg_pre_chess_rating, 0)
```


9. And bringing it all together now, in to a data frame which includes the required columns to answer the question above.

```{r, eval = TRUE}
score_df <- data.frame(player_name, state_name, points, pre_rating, avg_pre_chess_rating);
kable(score_df)
```

***
10. Writing the results in to a csv file

```{r, eval = TRUE}
write.csv(score_df, file ="SI_project1_results.csv")
```
